---
title: "Diabetes Classification"
author: "Lovedeep Bajaj"
date: "1/19/2021"
output: html_document
---

Features:   
__Age:__ how old the person is in years  
__Gender:__ Male or Female    
__Polyuria:__ No or Yes  
__Polydipsia:__ No or Yes  
__sudden.weight.loss:__ No or Yes  
__weakness:__ No or Yes  
__Polyphagia:__ No or Yes  
__Genital.thrush:__ No or Yes  
__visual.blurring Itching:__ No or Yes  
__Irritability:__ No or Yes  
__delayed.healing:__ No or Yes  
__partial.paresis:__ No or Yes  
__muscle.stiffness:__ No or Yes  
__Alopecia:__ No or Yes  
__Obesity:__ No or Yes    

Label:   
__class:__ Negative or Positive  

```{r}
# setting the working directory 
setwd("~/Documents/R/Projects/Diabetes Classification")

# read in the csv 
df <- read.csv('diabetes_data_upload.csv')

# viewing the first 5 values in the df
head(df)
```

```{r}
# seeing what the type of each column is 
sapply(df, class)
```

```{r}
# checking to see the length of the df 
nrow(df)
```

```{r}
# checking to see if any missing values in the features (there are no missing values) also making sure age 
# doesn't contain any outliers
summary(df)
```

```{r}
# checking to see how many duplicates are in the dataset 
nrow(df[duplicated(df), ])
```

## DATA VISUALIZATION
```{r}
# importing required library 
library('ggplot2')

# function for creating the plots
create.plots <- function(df, x, fill, title){
  plot.title = paste('Based On The', title, 'Feature Do People Have Diabetes')
  p <- ggplot(data = df, aes(x = {{x}}, fill = {{fill}})) + 
    geom_bar(position = position_dodge())+
    ggtitle(plot.title)+
    # removing the space between the x values and the bars
    scale_y_continuous(expand = c(0,0)) + scale_x_discrete(expand = c(0,0)) + 
    scale_fill_manual(values=c("#0a043c", "#03506f")) +
    theme_classic()+
    theme(
      plot.title = element_text(
        size = 12, # the size of the title
        face = 'bold', # making sure the title is bold
        hjust = 0.5, # centering title 
        vjust = 3 # moving the title up 
      ),
      axis.ticks.x = element_blank(), # removing the ticks on the x axis
      axis.title.y = element_blank(), # getting rid of the y axis title
      axis.ticks.y = element_blank(), # removing the ticks on the y axis
    )
  return(p)
}
```

### Age
```{r}
# binning the age
df$Age.bin <- cut(df$Age, breaks = c(9,19,29,39,49,59,69,79,89,99), 
                 labels = c('10-19', '20-29', '30-39', '40-49', '50-59', '60-69', '70-79', '80-89', '90-99'))
title = 'Age Group'

# creating the plot
create.plots(df,Age.bin, class, title)
```

### Gender
```{r}
title = 'Gender'
create.plots(df,Gender, class, title)
```

### Polyuria
```{r}
title = 'Polyuria'
create.plots(df,Polyuria, class, title)
```

### Polydipsia 
```{r}
title = 'Polydipsia'
create.plots(df,Polydipsia, class, title)
```

### Sudden weight loss 
```{r}
title = 'Sudden Weight Loss'
create.plots(df,sudden.weight.loss, class, title)
```

### Weakness 
```{r}
title = 'Weakness'
create.plots(df,weakness, class, title)
```

### Polyphagia 
```{r}
title = 'Polyphagia'
create.plots(df,Polyphagia, class, title)
```

### Genital thrush
```{r}
title = 'Genital Thrush'
create.plots(df,Genital.thrush, class, title)
```

### Visual blurring 
```{r}
title = 'Visual Blurring'
create.plots(df,visual.blurring, class, title)
```

### Itching   
```{r}
title = 'Itching'
create.plots(df,Itching, class, title)
```

### Irritability 
```{r}
title = 'Irritability'
create.plots(df,Irritability, class, title)
```

### Delayed healing
```{r}
title = 'Delayed Healing'
create.plots(df,delayed.healing, class, title)
```

### Partial paresis 
```{r}
title = 'Partial Paresis'
create.plots(df,partial.paresis, class, title)
```

### Muscle stiffness 
```{r}
title = 'Muscle Stiffness'
create.plots(df,muscle.stiffness, class, title)
```

### Alopecia 
```{r}
title = 'Alopecia'
create.plots(df,Alopecia, class, title)
```

### Obesity 
```{r}
title = 'Obesity'
create.plots(df,Obesity, class, title)
```

### Class
```{r}
title = 'How Many People Diabetes'
ggplot(data = df, aes(x = class)) + 
  geom_bar(fill = '#222831') +
  xlab('Class')+
  ggtitle(title) +
  # removing the space between the x values and the bars
  scale_y_continuous(expand = c(0,0)) + scale_x_discrete(expand = c(0,0)) + 
  theme_classic()+
  theme(
    plot.title = element_text(
      size = 12, # the size of the title
      face = 'bold', # making sure the title is bold
      hjust = 0.5, # centering title
      vjust = 3 # moving the title up
    ),
    axis.ticks.x = element_blank(), # removing the ticks on the x axis
    axis.title.y = element_blank(), # getting rid of the y axis title
    axis.ticks.y = element_blank(), # removing the ticks on the y axis
  )
```

## FEATURE TRANSFORMATION
```{r}
# importing required library
library(dplyr)

# converting all the features to type char
df %>%
  mutate_all(as.character) -> df
```

### Converting Gender
```{r}
# male = 0 and female = 1
df$Gender[df$Gender == 'Male'] <- 0
df$Gender[df$Gender == 'Female'] <- 1

# checking to make sure the change occurred 
unique(df$Gender)
```

### Converting Polyuria - Obesity
```{r}
# No = 0 and Yes = 1
df[df == 'No'] <- 0
df[df == 'Yes'] <- 1

# checking to see if the values of Polyuria changed 
unique(df$Polyuria)
```

### Converting Class
```{r}
# Negative = 0 and Positive = 1
df$class[df$class == 'Negative'] <- 0
df$class[df$class == 'Positive'] <- 1

# checking to see if the values of class changed 
unique(df$class)
```


```{r}
# function that does min-max scaling because I want the values to be in a range of 0 and 1 
normalize.min_max <- function(x){
  return((x - min(x))/(max(x) - min(x)))
}
```

### Scaling Age
```{r}
# converting Age to numeric type 
df$Age <- as.numeric(df$Age)

df$Age <- normalize.min_max(df$Age)

# checking the summary of Age to make sure the scaling worked
summary(df$Age)
```

### Converting Features to Factor
```{r}
# removing the Age.bin from the dataframe 
df <- subset(df, select=-c(Age.bin))

# converting the type of all the features to numeric
df %>%
  mutate_all(as.factor) -> df

# making sure all the values are numeric and that they all contain the correct values
summary(df)
```

### TRAINING AND TESTING SET
```{r}
# this preserves the results 
set.seed(101)

# import required library 
library(caTools)

# 70% training and 30% testing
split = sample.split(df$class, SplitRatio = 0.7)
training_df <- subset(df, split == TRUE)
testing_df <- subset(df, split == FALSE)
```

```{r}
# checking the number of rows in the training_df 
nrow(training_df)
```

```{r}
# checking to see how many people in the training_df have diabetes
nrow(training_df[training_df$class == 1,])
```

```{r}
# checking the number of rows in the testing_df 
nrow(testing_df)
```

```{r}
# checking to see how many people in the testing_df have diabetes
nrow(testing_df[testing_df$class == 1,])
```

## MODEL COMPARSION
Creating a decision tree and random forest model and seeing which one yields me the best results
```{r}
# importing required libraries
library(caret) # install.packages('caret', dependencies = TRUE) <- e1071 error will occur
library(party)

# decision tree model
tree.model <- ctree(class ~., data = training_df)

# predicting with the model 
tree.prediction <- predict(tree.model, testing_df)

# view the confusion matrix 
confusionMatrix(tree.prediction, testing_df$class)
```
This model is extremely good with an Accuracy of 92% and Sensitivity of .0.9333 Specificity : 0.9167. 


```{r}
# importing required library 
library(randomForest)

# random forest model
forest.model <- randomForest(class ~., data = training_df)

# predicting with the model 
forest.predict <- predict(forest.model, testing_df)

# viewing the confusion matrix 
confusionMatrix(forest.predict, testing_df$class)
```
This model gives us a better accuracy and specificity then the decision tree. The Accuracy is 96% and the Specificity is 1.0000. 

#### BEST MODEL: RANDOM FOREST




